I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-26 Paper/proofs.tex
b1 preparation

					\end{enumerate}
					That is, $a_k$ is not a source of a weak compensating pair. We denote 
					$A_{pure}$ by $I$ for short.
					\item $a_k \in A_{open}$ iff $a_k \notin A_{pure} \sqcup A_{zero}$, that is there exists two paths $\rho,\rho'$ from $a_k$, one with positive and one with negative weight. 	We denote $A_{open}$ by $J$ for short.
				\end{itemize}
			\end{definition}

@@

			Recall the sign function of nodes from Definition \ref{sign_of_nodes}. Because weak compensating paths are allowed, the sign function is no longer defined on all nodes. $S$ is defined on all nodes $a_i \in I$. 
			Further, we have:

			\begin{lemma}\label{lem:sign}
				For a node $n$, if there exists at least one path from a node $a_i\in I$ to $n$ with nonzero weight, then $S$ is defined on $n$.
			\end{lemma}

			\begin{proof}
				By contradiction, suppose that there exists two paths $\rho_1,\rho_2$ from $n$ to $z$ with (strictly) different signs. The pair $\langle a_i,n\rangle \cdot \rho_1$ and $\langle a_i,n\rangle \cdot \rho_2$ is compensating, a contradiction with $a_i\in I$. \qed
			\end{proof}

			Hence for any node $m$ such that $S$ is undefined on $m$, value of $m$ can only depends upon input nodes in $J$.

@@

			\begin{proof}
				By contradiction, assume that there exists two nodes $b,c$ such that $W_{bc}\neq 0$ and $c$ is in $D_J$ and $b \notin D_J$, then 
				there is a path $\rho$ from an input node $a_i \notin J$ to $b$
				 with nonzero weight. Then $\rho \cdot bc$ is a path from $a_i \notin J$ 
				 to $c$ with non zero weight, a contradiction. \qed
			\end{proof}

			By definition, for any $b\in D_J$, either $b\in B_{open}$, or $b$ is a constant node: any path from any input node must have zero weight to it.

			For an input $\vx$ and a subset $S$ of input nodes, we denote by $\boldsymbol{x}_S$ to refer the input vector $\langle x_{a_k}\rangle_{a_k\in S}$. We use $\boldsymbol{x}_I\oplus \boldsymbol{x}_J = \boldsymbol{x}$ and $\Val(z)=\Val_{\boldsymbol{x}}(z)=\Val_{\boldsymbol{x}_I,\boldsymbol{x}_J} (z)$. 
			For a node $b$ for which $S$ is undefined, $b$ only depends on $J$; so, if $\boldsymbol{x}_J$ is fixed, then the value of $b$ will be fixed.

			We define $\boldsymbol{x}_I^*$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=-1$). Similarly, we define $\boldsymbol{x}_I^\sharp$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=-1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=1$).

			\begin{lemma} \label{lem:reach_max_2}
				Let $b$ be an output node or a node in a hidden layers such that $S(b)=1$. Then for any valuation $\boldsymbol{x}^0_J$,	we have: 
				$$\max_{\boldsymbol{x}_I} (\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(b)) =  \Val_{\boldsymbol{x}_I^*,\boldsymbol{x}^0_J}(b).$$

				If $S(b)=-1$, then a similar results hold, replacing $\boldsymbol{x}^*_I$ by $\boldsymbol{x}^\sharp_I$. The opposite holds for $\min_{\boldsymbol{x}_I}$.
			\end{lemma}

			\begin{proof}

				For any fixed input vector $\boldsymbol{x}^0_J$, consider the DNN $D'$ with input nodes $a_i\in I$, with all $\boldsymbol{x}^0_J$ 
				%and their propagation in hidden layers 	
				as bias. We denote $\val'$ values of nodes In this DNN $D'$.
				 We have $\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(z) = \val'_{\vx}(\boldsymbol{x}_I)(z)$.

				In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply Lemma \ref{lem:key1} to get that $\Val_{x_I,\boldsymbol{x}^0_J}(z)$ reaches its maximal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$, and we are done. \qed
			\end{proof}

			From Lemma \ref{lem:reach_max_2}, we have the following corrolary:

@@

			$$\max(b)=\max_{\vx_j} \val_{\vx^{\star}_I,\vx_J}(b)$$

			and for $S(b)=-1$:

			$$\max(b)=\max_{\vx_j} \val_{\vx^{\sharp}_I,\vx_J}(b)$$
			\end{corollary}

>>>>

					\end{enumerate}
					Hence, $a_k$ is not a source of a weak compensating pair. We denote 
					$A_{pure}$ by $I$ for short.
					\item $a_k \in A_{open}$ iff $a_k \notin A_{pure} \sqcup A_{zero}$, that is there exists two paths $\rho,\rho'$ from $a_k$, one with positive and one with negative weight. 	We denote $A_{open}$ by $J$ for short.
				\end{itemize}
			\end{definition}

@@

			For an input $\vx$ and a subset $S$ of input nodes, we denote by $\boldsymbol{x}_S$ to refer the input vector $\langle x_{a_k}\rangle_{a_k\in S}$. We use $\boldsymbol{x}_I\oplus \boldsymbol{x}_J = \boldsymbol{x}$ and 
			$\Val_{\boldsymbol{x}_I,\boldsymbol{x}_J} (z)=\Val_{\boldsymbol{x}}(z)$.

			Invoking Lemma \ref{lem:subnetwork}, we can apply Theorem 1 on $D_I$, and obtain two input vectors $\boldsymbol{x}_I^*,\boldsymbol{x}_I^{\sharp}$ on $I$
			minimizing and maximizing all the nodes in $D_I$.

			Recall the sign function of nodes from Definition \ref{sign_of_nodes}. 
			Because weakly compensating paths are allowed, the sign function is no longer defined on all nodes. $S$ is defined for all nodes from which there is no weakly compensating pairs of paths to $z$. In particular, $S$ is defined on all nodes of $I$.

			\begin{lemma}\label{lem:sign}
				Let $m,n$ be two nodes, such that $S$ is defined on $m$ and there is a path $\rho$ with non zero weight to a node $n$. Then $S$ is defined on $n$
				 (there is no compensating pair of paths from $n$ to $z$).
				Further, $S(n)= Sign(weight(\rho)) S(m)$.
			\end{lemma}

			\begin{proof}
				By contradiction, assume that there exists two paths $\rho_1,\rho_2$ from 
				$n$ to $z$ with (strictly) different signs. The pair $\langle a_i,n\rangle \cdot \rho_1$ and $\langle a_i,n\rangle \cdot \rho_2$ is compensating, a contradiction with $a_i\in I$. \qed
			\end{proof}

			Hence, for all node $n$, either $S$ is defined on $n$, or 
			$n$ depends only upon input nodes in $J$.
			Let us call $D_J$ the set of such nodes which only depends upon nodes in $J$.
			$D_J$ forms a sub-network.
			$D_J$ contains the set of all nodes from which there is a weakly compensating pair of paths to $z$.
			By definition, for any $b\in D_J$, either $b\in B_{open}$, or $b$ is a constant node: any path from any input node must have zero weight to it.
			For a node $b$ for which $S$ is undefined, $b$ thus only depends on $J$; so, if $\boldsymbol{x}_J$ is fixed, then the value of $b$ will be fixed.
			Notice that $D_I \sqcup D_J$ does not cover the DNN entirely in general.

@@

			\begin{proof}
				By contradiction, assume that there exists two nodes $b,c$ such that $W_{bc}\neq 0$ and $c$ is in $D_J$ and $b \notin D_J$, then 
				there is a path $\rho$ from an input node $a_i \notin J$ to $b$
				 with nonzero weight. Then $\rho \cdot bc$ is a path from $a_i \notin J$ 
				 to $c$ with non zero weight, a contradiction. \qed
			\end{proof}
			\fi

			\begin{lemma} \label{lem:reach_max_2}
				Let $n$ be a node on which $S$ is defined. If $S(n)=1$,
				for any valuation $\boldsymbol{x}^0_J$,	we have: 
				$$\max_{\boldsymbol{x}_I} (\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(b)) =  \Val_{\boldsymbol{x}_I^*,\boldsymbol{x}^0_J}(b).$$

				$$\min_{\boldsymbol{x}_I} (\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(b)) =  \Val_{\boldsymbol{x}_I^{\sharp},\boldsymbol{x}^0_J}(b).$$

				If $S(b)=-1$, 
				$$\max_{\boldsymbol{x}_I} (\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(b)) =  \Val_{\boldsymbol{x}_I^{\sharp},\boldsymbol{x}^0_J}(b).$$

				$$\min_{\boldsymbol{x}_I} (\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(b)) =  \Val_{\boldsymbol{x}_I^{*},\boldsymbol{x}^0_J}(b).$$

			\begin{proof}	
				Fix an input vector $\boldsymbol{x}^0_J$.
				This also fixes the values in $D_J$.
				Consider the DNN $D'$ with input nodes $a_i\in I$, where all the 
				edges from $m \in D_J$ to $n$ are replaced by a bias (equals to the weight of the edge times the fix value $\val_{\boldsymbol{x}^0_J}(m)$) for $n$ (we summ all these bias for a neuron $n$).
				We denote $\val'$ values of nodes in this DNN $D'$.

				We have $\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(z) = \val'_{\boldsymbol{x}_I}(z)$.
				In the simplified $D'$, there is no weakly compensating path to $z$ because 
				$D_J$ contains all the source of weakly compensating path to $z$.
				Therefore we can apply Lemma \ref{lem:key1} to get that 
				$\val'_{x_I}(z)$ reaches its maximal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$, and conclude by using $\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(z) = \val'_{\boldsymbol{x}_I}(z)$. The case for min is similar. \qed
			\end{proof}

			From Lemma \ref{lem:reach_max_2}, we immediatly obtain the following corollary:

@@

			$$\max(b)=\max_{\vx_j} \val_{\vx^{\star}_I,\vx_J}(b)$$
			$$\min(b)=\max_{\vx_j} \val_{\vx^{\sharp}_I,\vx_J}(b)$$
			and for $S(b)=-1$:

			$$\max(b)=\max_{\vx_j} \val_{\vx^{\sharp}_I,\vx_J}(b)$$
			$$\min(b)=\max_{\vx_j} \val_{\vx^{\star}_I,\vx_J}(b)$$
			\end{corollary}
